EDA

### 2024-11-24 other/EDA.qmd
added per day graph

library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(scales)
library(knitr)
library(kableExtra)

df <- read_csv(here::here("data/analysis_data/augment_data.csv"))

@@

# Ensure 'award_date' is in Date format (if not already)
# df$award_date <- as.Date(df$award_date, format="%Y-%m-%d")  # Commented out since it's already a Date

# Identify the top 5 buyers by total amount awarded
top_buyers <- df %>%
  group_by(buyer) %>%
  summarise(total_amount = sum(amount, na.rm = TRUE)) %>%
  arrange(desc(total_amount)) %>%
  slice_head(n = 3) %>%
  pull(buyer)

# Filter the data for only the top 5 buyers
df_top_buyers <- df %>%
  filter(buyer %in% top_buyers)

# Create a 'year_month' column using lubridate
df_top_buyers <- df_top_buyers %>%
  mutate(year_month = floor_date(award_date, "month"))

# Aggregate the amount by 'year_month' and 'buyer'
monthly_spending <- df_top_buyers %>%
  group_by(year_month, buyer) %>%
  summarise(total_amount = sum(amount, na.rm = TRUE), .groups = 'drop')

# Plot the line graph
ggplot(monthly_spending, aes(x = year_month, y = total_amount, color = buyer, group = buyer)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Amount Spent Each Month by Top 5 Buyers",
       x = "Time",
       y = "Amount Awarded",
       color = "Buyer") +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

>>>>

library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(scales)
library(knitr)
library(kableExtra)
library(zoo)

df <- read_csv(here::here("data/analysis_data/cleaned_data.csv"))

@@

start_date_min <- min(df$start_date)
end_date_max <- as.Date("2030-12-31")

# Create events at start dates (add per_day amount)
events_start <- data.frame(date = df$start_date, amount = df$per_day)

# Create events at end dates + 1 day (subtract per_day amount)
events_end <- data.frame(date = df$end_date + 1, amount = -df$per_day)

# Combine start and end events
events <- rbind(events_start, events_end)

# Aggregate amounts by date
events <- aggregate(amount ~ date, data = events, sum)

# Sort events by date
events <- events[order(events$date), ]

# Compute cumulative sum of amounts to get the total per_day amount active on each date
events$cumulative_amount <- cumsum(events$amount)

# Create a full sequence of dates from the earliest start date to the latest end date
date_seq <- seq(start_date_min, end_date_max, by = "day")

# Merge cumulative amounts with the full date sequence
amounts <- data.frame(date = date_seq)
amounts <- merge(amounts, events[, c("date", "cumulative_amount")], by = "date", all.x = TRUE)

# Carry forward the last observed cumulative_amount to fill in missing dates
amounts$cumulative_amount <- na.locf(amounts$cumulative_amount, na.rm = FALSE)
amounts$cumulative_amount[is.na(amounts$cumulative_amount)] <- 0

# Plot the line chart using ggplot2
library(ggplot2)
ggplot(amounts, aes(x = date, y = cumulative_amount)) +
  geom_line(color = "blue") +
  labs(title = "Total Amount Spent Over Time",
       x = "Date",
       y = "Total Amount Spent per Day") +
  scale_y_continuous(labels = comma) +
  theme_minimal()
